---
title: "Examining the Influence of Urban Social Disorder Events on Deadly Incidents Across Geographic Regions"
author: "Emory Bayer"
institute: "American University"
date: "03/22/2024 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: 
      - "default"
      - "default-fonts"
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
    class: 
      - "themer-better-default" # Choose a professional theme
    dev: 'svg'  # Specify SVG as the graphics device
---               

```{r setup, include = FALSE} 
options(htmltools.dir.version = FALSE)
knitr :: opts_chunk$set(
  echo = FALSE,
  message = FALSE, 
  warning = FALSE ,
  fig.align = 'center')

library(tidyverse)
library(knitr)
library(cowplot)
library(patchwork)
library(ggplot2)
library(readxl)
library(dplyr)


setwd("C:/Users/emory/Documents/Statistics Level 2")
df = read_excel("Copy of cities.xlsx")
```
Class: inverse, bottom  

#Topic

###Data:
####This dataset, from the Urban Social Disorder Global Dataset, analyzes the impact of urbanization on global populations.

- covers 186 national capitals & urban centers from 1960 to 2014
- details social disorder events, both violent and non-violent
###Key Variables:
- DEATHEVENTS: social disorder events within a city which result in at least one death between 1960 and 2014
- NEVENTS: social disorder events within a city between 1960 and 2014
###Reason this dataset is important: 
- helps to understand if certain regions are more prone to experiencing violent social disorder outbreaks; essential for implementing mitigation

---
# Analysis 1.
<small>This visual analysis examines the relationship between DEATHEVENTS and city in the Middle East and North Africa (MENA)<small>

```{r, fig.width=6.5, fig.height=6.5}
Middle_E_NAfrica <- df %>% 
  filter(REGION == "Middle East and North Africa")

plot1 <- ggplot(Middle_E_NAfrica, aes(x = CITY, y = DEATHEVENTS)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black", size = 1) + 
  geom_text(aes(label = DEATHEVENTS), vjust = -0.5, color = "black", size = 3) +  
  labs(x = "City", y = "Number of Deadly Social Disorder Events", 
       title = "Deadly Social Disorder Events in MENA by City") +
  theme_minimal(base_size = 9) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "bold"),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size = 14, hjust = 0.5),
        axis.line = element_line(color = "black"),
        panel.grid.major = element_line(color = "gray"),
        panel.grid.minor = element_blank(),
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        plot.background = element_rect(fill = "white", color = "black"),
        panel.border = element_rect(color = "black", fill = NA),
        axis.ticks = element_line(color = "black", size = 0.5))

print(plot1)
```



---
# Analysis 2

  **Impact of number of events on the number of **
  **deadly events in North Africa and Latin America**
```{r, fig.width=8, fig.height=7}
Middle_E_NAfrica <- df %>% 
  filter(REGION == "Middle East and North Africa")

plot1 <- ggplot(Middle_E_NAfrica, aes(x = NEVENTS, y = DEATHEVENTS)) +
  geom_line(color = "purple") +
  geom_point() +
  scale_y_continuous(breaks = seq(0, max(Middle_E_NAfrica$DEATHEVENTS), by = 50)) +
  scale_x_continuous(breaks = seq(0, max(Middle_E_NAfrica$NEVENTS), by = 100)) +
  labs(x = "Number of events", y = "Number of Deadly events") +  
  annotate("text", x = Inf, y = Inf, label = "MENA", size = 6, hjust = 1.6, vjust = 1) +
  theme(axis.title.y = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12, face = "bold"),
        axis.title.x = element_text(margin = margin(t = 15), face = "bold"),
        panel.border = element_rect(color = "black", fill = NA))


Latin_America <- df %>%
  filter(REGION == "Latin America")


plot4 <- ggplot(Latin_America, aes(x = NEVENTS, y = DEATHEVENTS)) +
  geom_line(color = "red") +
  geom_point() +
  scale_y_continuous(breaks = NULL, limits = c(0, 350), minor_breaks = seq(0, 350, by = 50)) +
  scale_x_continuous(breaks = seq(0, max(Latin_America$NEVENTS), by = 50)) +
  labs(x = "Number of events", y = NULL) +  
  theme(panel.grid.major = element_line(color = "white", linetype = "solid"),
        panel.grid.minor = element_line(color = "white", linetype = "solid"),
        axis.text = element_text(size = 12, face = "bold"),
        axis.title.x = element_text(margin = margin(t = 15), face ="bold"), 
        axis.title.y = element_blank(),
        panel.border = element_rect(color = "black", fill = NA)) +
  annotate("text", x = Inf, y = Inf, label = "Latin America", size = 6, hjust = 1.1, vjust = 1)


combined_plots <- plot1 + plot4 +
  plot_layout(ncol = 2) +  
  theme(
    plot.margin = margin(40, 40, 40, 40),  
    axis.title = element_text(margin = margin(t = 20))
  )

print(combined_plots)
```

